Technique for \earlywarning\ detection of compact binary coalescence and its implications for multi-messenger astronomy

### 2011-06-10 method.tex
Fixing column overflow for long equation
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1007 ccdd05d5-56c1-461a-bf98-85f6876523b4
Reorganize section 2.1 to make grumpy grad students happy.
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1008 ccdd05d5-56c1-461a-bf98-85f6876523b4

Inspiral signals are continuously parameterized by a set of intrinsic source
parameters $\mathbf\Theta$ that determine the amplitude and phase evolution of the
\GW{} strain. For systems where the effects of spin can be
ignored, the intrinsic source parameters are the component masses of the
binary, $\mathbf\Theta = (m_1, m_2)$. Searches for inspiral signals typically
employ matched filter banks that discretely sample the possible intrinsic
parameters~\citep{findchirppaper}.  For a given source, the strain observed by
the detector is a linear combination of two waveforms corresponding to the
`$+$' and `$\times$' \GW{} polarizations.  Thus, for any value of
$\mathbf\Theta$ we must implement 2 filters.  The coefficients for the $\numtmps$ filters
are known as templates, and are formed by discretizing and time reversing the
waveforms and weighting them by the inverse amplitude spectral density of the
detector's noise. To construct a template bank, templates are chosen with the
$\numtmps/2$ discrete signal parameters $\mathbf\Theta_0,\, \mathbf\Theta_1,\, \dots,\,
\mathbf\Theta_{\numtmps/2-1}$ to assure a bounded loss of \SNR~\citep{Owen:1998dk}. That
is, any possible signal within a given range of intrinsic
source parameters will have an inner product that is $\geqslant 0.97$ with at
least one template. Such a template bank is said to have a {\em minimum match}
of 0.97. The data from the detector are whitened and convolved with each
template to produce $\numtmps$ \SNR\ time series. Local peak-finding across time and
templates determines detection candidates.

Filtering the detector data involves a convolution of the data with the
templates.  For a unit-normalized template $h_i[k]$ and whitened detector data
$x[k]$, both sampled at a rate $f^0$, the result can be interpreted as the
\SNR{}, $\rho_i[k]$ defined as
%
% Filtering equations
%
\begin{equation}
	\label{eq:SNRTD}
	\rho_i [k] = \sum_{n=0}^{N-1} h_{i}[n] x [k-n].
\end{equation}

@@

From the combination of transformations to the compact binary templates defined
in equation \eqref{eq:time-sliced-templates} and \eqref{eq:svddecomp} we define
the early-warning filter output accumulated up to time slice $s$ as,
%
% orthogonal decomposition filtering
%
\begin{equation}
	\rho_i^s [k] =%
		% Reconstruction
		\underbrace{
			\sum_{\mathclap{l=0}}^{\mathclap{L^s-1}} v_{il}^s \sigma_l^s
		}_\textrm{\clap{reconstruction}}
		% Orthogonal FIR filter
		\overbrace{
			\sum_{\mathclap{n=0}}^{\mathclap{N^s-1}} u_l^s[n] x^s[k-n]
		}^\textrm{\clap{orthogonal {\sc fir} filters}}
		% Plus ...
		+
		% Interpolation SNR
		\underbrace{
			\left(H^\uparrow \rho_i^{s+1}\right)[k]
		}_\textrm{\clap{{\sc snr} from previous time slices}}
\end{equation}
%
%
This quantity is related to the early-warning \SNR{} by the normalization
of the partial filter defined by the present time slice, which may be computed
in advance. \editorial{Chad: are we happy with that statement?}
%
%
\begin{figure*}[h!]
	\begin{center}
		\includegraphics{figures/lloid-diagram.pdf}
		\caption{\label{fig:pipeline} Schematic of \lloid{} pipeline illustrating
signal flow.  Circles with arrows represent interpolation
\protect\includegraphics{figures/upsample-symbol.pdf} or decimation
\protect\includegraphics{figures/downsample-symbol.pdf}.  Circles with plus
signs represent summing junctions
\protect\includegraphics{figures/adder-symbol.pdf}.  Squares
\protect\includegraphics{figures/fir-symbol.pdf} stand for \fir{} filters.  Sample
rate decreases from the top of the diagram to the bottom.  In this diagram each
time slice contains three \fir\ filters that are linearly combined to produce
four output channels.  In a typical pipeline the number of \fir\ filters is
much less than the number of output channels.}
	\end{center}
\end{figure*}
%
%
The signal flow diagram in figure~\ref{fig:pipeline} illustrates how this
recursion relation can be realized as a multi-rate filter network with outputs for
each of the early-warning \SNR{}s.

>>>>

Inspiral signals are continuously parameterized by a set of intrinsic source
parameters $\mathbf\Theta$ that determine the amplitude and phase evolution of the
\GW{} strain. For systems in which the effects of spin can be ignored, the intrinsic
source parameters are the just component masses of the binary,
 $\mathbf\Theta = (m_1, m_2)$. For a given source, the strain observed by the
 detector is a linear combination of two waveforms corresponding to the
`$+$' and `$\times$' \GW{} polarizations.  Thus, for any value of
$\mathbf\Theta$ we must implement two filters (real-valued) filters.

Searches for inspiral signals typically employ matched filter
banks that discretely sample the possible intrinsic parameters~\citep{findchirppaper}.
To construct a template bank, templates are chosen with
$\numtmps/2$ discrete signal parameters $\mathbf\Theta_0,\, \mathbf\Theta_1,\, \dots,\,
\mathbf\Theta_{\numtmps/2-1}$ to assure a bounded loss of \SNR~ for signals given by
intermediate values of the intrinsic parameters~\citep{Owen:1998dk}. That
is, any possible signal within a given range of intrinsic
source parameters will have an inner product that is $\geqslant 0.97$ with at
least one template. Such a template bank is said to have a {\em minimum match}
of 0.97.

Filtering the detector data involves a convolution of the data with the
templates.  For a unit-normalized template $h_i[k]$ and whitened detector data
$x[k]$, both sampled at a rate $f^0$, the result can be interpreted as the
\SNR{}, $\rho_i[k]$ defined as
%
% Filtering equations
%
\begin{equation}
	\label{eq:SNRTD}
	\rho_i [k] = \sum_{n=0}^{N-1} h_{i}[n] x [k-n].
\end{equation}
The coefficients for the $\numtmps$ filters are known as templates, 
and are formed by discretizing and time reversing the
waveforms and weighting them by the inverse amplitude spectral density of the
detector's noise. The data from the detector are whitened and convolved with each
template to produce $\numtmps$ \SNR\ time series. Local peak-finding across time and
templates determines detection candidates.

@@

From the combination of transformations to the compact binary templates defined
in equation \eqref{eq:time-sliced-templates} and \eqref{eq:svddecomp} we define
the early-warning filter output accumulated up to time slice $s$ as,
%
% orthogonal decomposition filtering
%
\begin{multline}
	\rho_i^s [k] =%
		% Interpolation SNR
		\overbrace{
			\left(H^\uparrow \rho_i^{s+1}\right)[k]
		}^\textrm{\clap{{\sc snr} from previous time slices}} \\
		% Plus ...
		+
		% Reconstruction
		\underbrace{
			\sum_{\mathclap{l=0}}^{\mathclap{L^s-1}} v_{il}^s \sigma_l^s
		}_\textrm{\clap{reconstruction}}
		% Orthogonal FIR filter
		\overbrace{
			\sum_{\mathclap{n=0}}^{\mathclap{N^s-1}} u_l^s[n] x^s[k-n]
		}^\textrm{\clap{orthogonal {\sc fir} filters}} .
\end{multline}
%
%
This quantity is related to the early-warning \SNR{} by the normalization
of the partial filter defined by the present time slice, which may be computed
in advance. \editorial{Chad: are we happy with that statement?}
%
%
\begin{figure*}[h!]
	\begin{center}
		\includegraphics{figures/lloid-diagram.pdf}
		\caption{\label{fig:pipeline} Schematic of \lloid{} pipeline illustrating
signal flow.  Circles with arrows represent interpolation
\protect\includegraphics{figures/upsample-symbol.pdf} or decimation
\protect\includegraphics{figures/downsample-symbol.pdf}.  Circles with plus
signs represent summing junctions
\protect\includegraphics{figures/adder-symbol.pdf}.  Squares
\protect\includegraphics{figures/fir-symbol.pdf} stand for \fir{} filters.  Sample
rate decreases from the top of the diagram to the bottom.  In this diagram each
time slice contains three \fir\ filters that are linearly combined to produce
four output channels.  In a typical pipeline the number of \fir\ filters is
much less than the number of output channels.}
	\end{center}
\end{figure*}
%
%
The signal flow diagram in figure~\ref{fig:pipeline} illustrates how this
recursion relation can be realized as a multi-rate filter network with outputs for
each of the early-warning \SNR{}s.
